---
title: "Identify scrambles that significant change expression with t-test"
output: html_notebook
---

First, let's read in the barcode counts and process that data.

```{r, echo = F}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(cowplot)
library(plotrix)
options(stringsAsFactors = F)
options(scipen = 10000)
```

```{r}
args <- c('../../processed_data/endo_scramble',
          '../../processed_data/endo_scramble/endo_scramble_combined_bc_map.txt',
          '../../processed_data/endo_scramble/endo_scramble_combined_controls_bc_map.txt',
          '../../processed_data/endo_scramble/endo_scramble_expression_formatted_std.txt')
count_folder <- args[1]
bc_map_file <- args[2]
bc_map_file_controls <- args[3]
formatted_file <- args[4]

# read in barcode count files
filelist = list.files(path = count_folder,
                      pattern = '^counts_*',
                      full.names = T)

# should be named so files are in following order, 1st number is biological replicate
# and second number is technical replicate
sample_names <- c('DNA1_1', 'DNA1_2', 'DNA2_1', 'DNA2_2', 'RNA1_1', 'RNA1_2', 'RNA2_1', 'RNA2_2')

for(i in seq(1:length(filelist))) {
    sample_name <- sample_names[i]
    x <- read.table(filelist[i], col.names=c(sample_name, 'barcode'), header = F)
    x[[sample_name]] <- 1000000*x[[sample_name]]/sum(x[[sample_name]])  #Normalizes by RPM
    assign(sample_name, x)  
}

count_list <- list(DNA1_1, DNA1_2, DNA2_1, DNA2_2, RNA1_1, RNA1_2, RNA2_1, RNA2_2)
 
# combine reads for all barcodes 
# feed list into reduce to full join entire list of data frames
all_counts <- count_list %>% purrr::reduce(full_join, by = "barcode")
print(paste("Number of unique sequenced barcodes, all replicates in DNA and RNA:", nrow(all_counts)))

# rearrange columns
all_counts <- select(all_counts, barcode, DNA1_1, DNA1_2:RNA2_2)
# remove individual sample files
rm(count_list, x)
rm(list = sample_names)

# read in mapping files
bc_map <- read.table(bc_map_file, header = F, sep = ',', 
                       col.names = c('barcode', 'variant', 'count'))

# read in formatted file so we can label mapping file
data <- read.table(formatted_file, header = T)

bc_map <- bc_map %>% 
    left_join(select(data, name, variant), by = 'variant')

# add in positive controls
bc_map_controls <- read.table(bc_map_file_controls, header = T, 
                           col.names = c('barcode', 'num_unique_variant', 'count',
                                         'count_most_common', 'variant', 'name'))
bc_map <- bind_rows(bc_map, select(bc_map_controls, barcode, variant, count, name))
print(paste("Number of barcodes in mapping run:", nrow(bc_map)))

# only keep those that match reference
bc_map <- filter(bc_map, !is.na(name))
print(paste("Number of barcodes in mapping run and mapped to reference:", nrow(bc_map)))

# only keep barcode counts for mapped barcodes
mapped_bc_counts <- inner_join(bc_map, all_counts, by = 'barcode')
# replace NA with 0
mapped_bc_counts[is.na(mapped_bc_counts)] <- 0
print(paste("Number of mapped and sequenced barcodes, all replicates in DNA and RNA:", nrow(mapped_bc_counts)))
```

We want to test if the difference in group means between the unscrambled and scrambled barcodes is
significantly different. We'll do this at the barcode level, where each barcode expression measurement
(DNA/RNA) in each of the four replicates is considered an individual observation. So, if a barcode is present
in all four replicates, there would be four independent observations for the t-test.

```{r}
mapped_bc_counts <- left_join(mapped_bc_counts, 
                              select(data, tss_name:scramble_pos_rel_tss,
                                     variant, category), by = 'variant')

mapped_bc_counts_melt <- mapped_bc_counts %>% 
    mutate(expn_1_1 = RNA1_1 / DNA1_1,
        expn_1_2 = RNA1_2 / DNA1_2,
        expn_2_1 = RNA2_1 / DNA2_1,
        expn_2_2 = RNA2_2 / DNA2_2) %>% 
    select(name, tss_name, category, expn_1_1:expn_2_2) %>% 
    reshape2::melt(id = c('name', 'tss_name', 'category'), value.name = 'expn')

scrambled_bcs <- mapped_bc_counts_melt %>% 
    filter(category == 'scramble')

unscrambled_bcs <- mapped_bc_counts_melt %>% 
    filter(category == 'unscrambled')
```

```{r, echo = F}
ttest_custom <- function(df, df_wt) {
    wt_name <- df$tss_name[1]
    # bind unscrambled barcodes to df
    df_with_wt <- bind_rows(df,
                            filter(df_wt, tss_name == wt_name)) %>% 
        mutate(category_fctr = factor(category)) %>% 
        filter(!is.na(expn), is.finite(expn))
    result <- tryCatch(
        {
            t.test(expn ~ category_fctr, df_with_wt)
        }, warning = function(cond) {
            return(NA)
        }, error = function(cond) {
            return(NA)
        }
    )
    return(result)
}

# test <- filter(scrambled_bcs, name == 'TSS_10266_regulondb,2723831,-_scrambled5-15_flipped_rc')

scramble_ttests <- scrambled_bcs %>% 
    group_by(name, tss_name) %>% 
    do(ttest = ttest_custom(., df_wt = unscrambled_bcs)) %>% 
    broom::tidy(ttest) %>% 
    ungroup()
```

```{r}
scramble_ttests <- scramble_ttests %>% 
    select(name, tss_name, mean_diff = estimate, mean_scramble = estimate1,
           mean_unscrambled = estimate2, tstat = statistic, p.value,
           conf.low, conf.high)

scramble_ttests %>% 
    filter(is.na(mean_diff)) %>% nrow()
```

```{r}
# adjust p-value
# n <- nrow(filter(scramble_ttests, !is.na(mean_diff)))
scramble_ttests$p.value.fdr <- p.adjust(scramble_ttests$p.value, method = 'fdr')

scramble_ttests <- scramble_ttests %>% 
    mutate(site_type_ttest = case_when(.$mean_diff < 0 ~ 'reduced',
                                 .$mean_diff > 0 ~ 'increased'),
           significant = ifelse(p.value.fdr <= 0.01, T, F))

# format name
scramble_ttests <- scramble_ttests %>% 
    mutate(name = gsub('_rc', '', name))
```

```{r}
# write results
write.table(scramble_ttests, file = '../../processed_data/endo_scramble/endo_scramble_ttests.txt',
            row.names = F, quote = F, sep = '\t')
```

Merge significant scrambles, by type (increased/reduced)

```{r}
scramble_ttests %>% 
    filter(significant == T, site_type_ttest == 'increased') %>% 
    mutate(chrom = 'U00096.2') %>% 
    left_join(., select(data, name, scramble_start_pos, scramble_end_pos), by = 'name') %>% 
    filter(!is.na(scramble_start_pos)) %>% 
    select(chrom, scramble_start_pos, scramble_end_pos, name, mean_diff, site_type_ttest) %>% 
    arrange(scramble_start_pos) %>% 
    write.table( '../../processed_data/endo_scramble/scrambles_ttest_increased.bed',
            sep = '\t', col.names = F, row.names = F, quote = F)

scramble_ttests %>% 
    filter(significant == T, site_type_ttest == 'reduced') %>% 
    mutate(chrom = 'U00096.2') %>% 
    left_join(., select(data, name, scramble_start_pos, scramble_end_pos), by = 'name') %>% 
    filter(!is.na(scramble_start_pos)) %>% 
    select(chrom, scramble_start_pos, scramble_end_pos, name, mean_diff, site_type_ttest) %>% 
    arrange(scramble_start_pos) %>% 
    write.table( '../../processed_data/endo_scramble/scrambles_ttest_reduced.bed',
            sep = '\t', col.names = F, row.names = F, quote = F)
```

```{bash}
bedtools merge -i ../../processed_data/endo_scramble/scrambles_ttest_increased.bed -c 5,4,4 -o mean,distinct,count_distinct > ../../processed_data/endo_scramble/scrambles_ttest_increased_merged.bed

bedtools merge -i ../../processed_data/endo_scramble/scrambles_ttest_reduced.bed -c 5,4,4 -o mean,distinct,count_distinct > ../../processed_data/endo_scramble/scrambles_ttest_reduced_merged.bed

```


```{r}
scramble_merged <- bind_rows(
    read.table('../../processed_data/endo_scramble/scrambles_ttest_increased_merged.bed',
               header = F, sep = '\t', 
               col.names = c('chrom', 'scramble_start_pos', 'scramble_end_pos',
                             'mean_group_diff', 'scrambles', 'num_scrambles')) %>% mutate(site_type = 'increased'),
    read.table('../../processed_data/endo_scramble/scrambles_ttest_reduced_merged.bed',
               header = F, sep = '\t', 
               col.names = c('chrom', 'scramble_start_pos', 'scramble_end_pos',
                             'mean_group_diff', 'scrambles', 'num_scrambles')) %>% mutate(site_type = 'reduced'))

table(scramble_merged$site_type)
```

Overlap TF sites with merged sites

```{r}
# http://regulondb.ccg.unam.mx/menu/download/datasets/files/BindingSiteSet.txt
tf_sites_all <- read.table('../../ref/regulondb_tfbs_U00096.2.txt', comment.char = '#',
                       header = F, sep = '\t',
                       col.names = c('tf_id', 'tf_name', 'tfbs_id', 'tfbs_left',
                                     'tfbs_right', 'strand', 'tf_gene_id', 'tx_unit',
                                     'expression_effect', 'promoter_name',
                                     'center_pos_relative_tss', 'tfbs_sequence',
                                     'evidence'))

# extract TFBS from sequence, in upper case
# grab upper case part of site corresponding to binding site
extract_upper <- function(string, toString) {
    replace_lower <- strsplit(string, "[[:lower:]]*")[[1]]
    only_upper <- replace_lower[replace_lower != ""]
    if(toString == T) {
        return(paste(only_upper, collapse = ''))
    }
    else{
        return(only_upper)
    }
}

tf_sites_all$tfbs <- unlist(lapply(tf_sites_all$tfbs_sequence, extract_upper, toString = T))

tf_sites <- tf_sites_all %>% 
    distinct(tf_name, tfbs_left, tfbs_right, expression_effect) %>% 
    filter(tfbs_left != 0) %>% 
    arrange(tfbs_left)

tf_sites %>% 
    mutate(chrom = 'U00096.2') %>% 
    select(chrom, tfbs_left, tfbs_right, tf_name) %>% 
    write.table('../../processed_data/endo_scramble/tf_sites.bed', 
                sep = '\t', col.names = F, row.names = F, quote = F)
```

```{bash}
# require 10% of TFBS to overlap with merged site, report overlap or NULL
cut -f 1,2,3,5 ../../processed_data/endo_scramble/scrambles_ttest_increased_merged.bed | bedtools intersect -a stdin -b ../../processed_data/endo_scramble/tf_sites.bed -loj -F 0.10 > ../../processed_data/endo_scramble/scrambles_ttest_increased_merged_tf_overlap10.bed

cut -f 1,2,3,5 ../../processed_data/endo_scramble/scrambles_ttest_reduced_merged.bed | bedtools intersect -a stdin -b ../../processed_data/endo_scramble/tf_sites.bed -loj -F 0.10 > ../../processed_data/endo_scramble/scrambles_ttest_reduced_merged_tf_overlap10.bed
```

```{r}
merged_overlap <- bind_rows(
    read.table('../../processed_data/endo_scramble/scrambles_ttest_increased_merged_tf_overlap10.bed',
                             header = F, fill = T,
                             col.names = c('chrom', 'scramble_start_pos', 'scramble_end_pos', 'scrambles',
                                           'chrom_tf', 'tfbs_left', 'tfbs_right', 'tf_name')) %>% 
    select(-chrom, -chrom_tf) %>% 
    mutate(site_type = 'increased'),
    read.table('../../processed_data/endo_scramble/scrambles_ttest_reduced_merged_tf_overlap10.bed',
                             header = F, fill = T,
                             col.names = c('chrom', 'scramble_start_pos', 'scramble_end_pos', 'scrambles',
                                           'chrom_tf', 'tfbs_left', 'tfbs_right', 'tf_name')) %>% 
    select(-chrom, -chrom_tf) %>% 
    mutate(site_type = 'reduced')) %>% 
    mutate(tf_name = ifelse(tf_name == '.', NA, tf_name)) %>% 
    distinct()
```

```{r}
scramble_merged <- scramble_merged %>% 
    left_join(select(merged_overlap, -scramble_start_pos, 
                     -scramble_end_pos, -site_type), 
              by = 'scrambles') %>% 
    left_join(select(tf_sites, tfbs_left, tfbs_right, tf_name, 
                     tf_effect = expression_effect) %>% distinct(),
              by = c('tfbs_left', 'tfbs_right', 'tf_name'))

# merged site agrees if any of the component scrambles agreed with TFBS effect
scramble_merged_parts <- scramble_merged %>% 
    mutate(scramble_list = strsplit(scrambles, split = ',TSS')) %>% 
    unnest(scramble_list) %>% 
    mutate(scramble_part = gsub('TSS', '', scramble_list)) %>% 
    mutate(scramble_part = paste0('TSS', scramble_part)) %>% 
    mutate(scramble_part = gsub('TSSneg', 'neg', scramble_part)) %>% 
    select(-scramble_list) %>% 
    left_join(select(data, name, tss_name, relative_exp, unscrambled_exp),
              by = c('scramble_part' = 'name'))
    
merged_tf_concordance <- function(scramble_name, tf_effect, scramble_merged_parts) {
    scramble_parts <- filter(scramble_merged_parts, scrambles == scramble_name)
    if(is.na(tf_effect)) {
        concordance <- NA
    } else if (tf_effect == '-') {
        concordance <- any(scramble_parts$mean_group_diff > 0)
    } else if (tf_effect == '+') {
        concordance <- any(scramble_parts$mean_group_diff < 0)
    } else if (tf_effect == '+-') {
        concordance <- TRUE
    } else {
        concordance <- NA
    }
    return(concordance)
}

scramble_merged$tf_concordance <- mapply(merged_tf_concordance,
                                         scramble_merged$scrambles,
                                         scramble_merged$tf_effect,
                                         MoreArgs = list(scramble_merged_parts = scramble_merged_parts))

write.table(scramble_merged, "../../processed_data/endo_scramble/scramble_merged_sites.txt",
            sep = '\t', row.names = F, quote = F)
```


