---
title: "Figure1_TSSscreen"
author: "Guillaume Urtecho"
date: "4/1/2019"
output: html_document
---

```{r setup, include=FALSE}
# setwd("~/Documents/projects/ecoli_promoters/endo/scripts/figs")
# knitr::opts_knit$set(root.dir = '~/Google Drive/Kosuri_Lab/Endo_Paper/scripts')

library(dplyr)
library(ggplot2)
require(cowplot)
library(ggsignif)
library(Biostrings)
library(stringr)

options(stringsAsFactors = F)

Endo2 <- read.table("../../processed_data/endo_tss/lb/rLP5_Endo2_lb_expression_formatted_std.txt",
                    header = T, fill  = T)

ref <- read.table('../../ref/endo_lib_2016_controls_clean.txt',
                  col.names = c('name', 'sequence'))
```

Figure1A - Venn Diagram Values

```{r}

Storz <- Endo2[grepl('storz', Endo2$name),] #11899
Wanner <- Endo2[grepl('wanner', Endo2$name),] #991
regulondb <- Endo2[grepl('regulondb', Endo2$name),] #4907


storz_wanner <- inner_join(Storz, Wanner, by = 'name')
wanner_regulondb <- inner_join(Wanner, regulondb, by = 'name')
regulondb_storz <- inner_join(regulondb, Storz, by = 'name')
all <- regulondb_storz[grepl('wanner', regulondb_storz$name),]

regulondb_storz <- anti_join(regulondb_storz, all, by = 'name')
wanner_regulondb <- anti_join(wanner_regulondb, all, by = 'name')
storz_wanner <- anti_join(storz_wanner, all, by = 'name')

#unique to each group
Storz <- anti_join(Storz, all, by = 'name') %>%
    anti_join(., regulondb_storz, by = 'name') %>%
    anti_join(., storz_wanner, by = 'name')

Wanner <- anti_join(Wanner, all, by = 'name') %>%
    anti_join(., wanner_regulondb, by = 'name') %>%
    anti_join(., storz_wanner, by = 'name')

regulondb <- anti_join(regulondb, all, by = 'name') %>%
    anti_join(., wanner_regulondb, by = 'name') %>%
    anti_join(., regulondb_storz, by = 'name')

#total number unique to each group
nrow(Storz) %>% paste(., "Unique to Thomason")#11032
nrow(Wanner) %>% paste(., "Unique to Conway")# 766
nrow(regulondb) %>% paste(., "Unique to RegulonDB")
nrow(storz_wanner) %>% paste(., "shared by Thomason and Conway")#111
nrow(wanner_regulondb) %>% paste(., "shared by wanner and RegulonDB")#41
nrow(regulondb_storz) %>% paste(., "shared by regulondb and Thomason")#683
nrow(all) %>% paste(., "share by all")#73
```

Figure 1C - Technical replicates

```{r}
neg <- subset(Endo2, grepl("neg_control", Endo2$name))
# this has been standardized already to be equal to 1
threshold <- 1

cor <- cor(Endo2$RNA_exp_1, Endo2$RNA_exp_2)

ggplot(Endo2, aes(RNA_exp_1, RNA_exp_2)) + 
  geom_point() +
  scale_x_log10() + scale_y_log10() + annotation_logticks(sides= 'bl') +
  annotate("text", label = paste("italic(r)==", signif(cor, 3)), parse = TRUE,
           x = 0.5, y = 15, size = 8) +
  xlab('Technical Replicate #1') + ylab('Technical Replicate #2') +
  ggtitle('Correlation between replicate TSS measurements') +
  geom_point(data=neg,aes(RNA_exp_1, RNA_exp_2), color = "firebrick1") +
  annotate("text", label = "Negative Controls", x = 20, y = .5, color = 'firebrick1', size = 8) +
  theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))

ggsave('../../figs/fig1/tss_replicate_corr.pdf', units = 'in', width = 7.29, height = 4.51)
```


Figure 1D, distributions of promoter activity and expression threshold

```{r}
rrsE<- c('TSS_16491_storz,4205882,+', 'TSS_16492_regulondb,4205994,+', 'TSS_16493_regulondb,4206055,+')

PrrsE <- Endo2$expn_med_fitted_scaled[Endo2$name %in% rrsE]
mid_PrrsE <- mean(PrrsE)
rm(rrsE)

ggplot(Endo2, aes(expn_med_fitted_scaled)) +
      geom_histogram(color = 'blue', fill = 'blue', alpha = .2, bins = 100) +
      geom_histogram(data=neg, aes(expn_med_fitted_scaled), color = 'red', fill = 'red', alpha = .2, bins = 100) +
      scale_x_log10(limits = c(0.35,100)) +
      scale_y_log10() +  
      annotation_logticks(sides = 'bl') +
      annotate("text", x= 4, y = 3000, label = "Active Threshold", size = 8) +
      geom_vline(xintercept = threshold, linetype = 'dashed') +
      geom_point(data=as.data.frame(PrrsE), aes(y = 10, x = PrrsE), shape = 25) + 
      annotate('text', x = mid_PrrsE, y = 16, label = 'PrrsE', fontface = 'italic', size = 6) +
       labs(x = 'rLP5 Expression', y = 'Count', title = 'Identifying Active Promoter Threshold') +
      theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))

ggsave('../../figs/fig1/tss_lib_expression_dist.pdf', units = 'in', width = 7.29, height = 4.51)
```

Figure 1E - Number of active promoters per source
```{r}
data.frame(Expression = c("Active TSS", "Inactive TSS"),
                         Value = c(nrow(filter(Wanner, expn_med_fitted_scaled > threshold)),
                                   nrow(filter(Wanner, expn_med_fitted_scaled < threshold))),
                         source = "Wanner") %>%
  rbind(., data.frame(Expression = c("Active TSS", "Inactive TSS"),
                         Value = c(nrow(filter(Storz, expn_med_fitted_scaled > threshold)),
                                   nrow(filter(Storz, expn_med_fitted_scaled < threshold))),
                         source = "Storz")) %>%
 rbind(., data.frame(Expression = c("Active TSS", "Inactive TSS"),
                         Value = c(nrow(filter(regulondb, expn_med_fitted_scaled > threshold)),
                                   nrow(filter(regulondb, expn_med_fitted_scaled < threshold))),
                         source = "RegulonDB")) %>%
        ggplot(aes(x=source, y=Value, fill = Expression)) + 
          geom_bar(stat="identity", position = position_dodge(width = 0), color = 'black') +
          scale_fill_manual(values = c("#2E6EB7", "#F6BA60")) +
          scale_x_discrete(expand = c(0, 2)) +
          theme(legend.position = c(0,1),
                axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20)) +
          labs(x="Source", y="Count")

ggsave('../../figs/fig1/activity_by_source.pdf', units = 'in', height = 8.1, width = 7.46)
```

Alternative to Fig1E, general proportion of active/inactive, not broken down by source.

```{r}
ggplot(Endo2, aes(active)) + 
    geom_bar(aes(fill = active)) +
    scale_fill_manual(values = c("#2E6EB7", "#F6BA60")) +
    scale_x_discrete(expand = c(0, 2)) +
    theme(legend.position = c(0,1),
          axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
          axis.title.x = element_text(size = 20),
          axis.text.y = element_text(size = 20), 
          axis.title.y = element_text(size = 20), 
          legend.text=element_text(size=20)) 

ggsave('../../figs/fig1/active_bar.pdf', units = 'in', height = 5, width = 4)
```



Figure 1F - inactive promoters are enriched for -10 sequences but not -35s
```{r}
# join sequences to expression data
Endo2_stats <- Endo2 %>%
            mutate("classification" = ifelse(expn_med_fitted_scaled > threshold,
                                             "Active", "Inactive"))
   
Endo2_stats[grepl(pattern = 'neg_control', x=Endo2_stats$name),]$classification <- "Negative \n Controls"
```


Figure 1E cont. - Score -10 motifs 
```{r echo=T, warning=FALSE, results='hide'}
#This code uses the pwms from bTSSFinder (which come from literature) to find the best -10 site
A = c(0.0097,	1.0000,	0.3363,	0.5335,	0.4963,	0.0781)
C = c(0.0618,	0.0000,	0.1190,	0.1094,	0.2299,	0.0268)
G = c(0.1042,	0.0000,	0.0856,	0.1317,	0.1399,	0.0000)
T = c(0.8244,	0.0000,	0.4591,	0.2254,	0.1339,	0.8951)
pwm_s70 <- data.frame(A,C,G,T)
pwm_s70 <- t(pwm_s70)

Endo2_stats$minus10_matches <- sapply(as.character(Endo2_stats$variant),
                                      function(x) matchPWM(pwm_s70,
                                                           str_sub(x,-150,-1)))

Endo2_stats$minus10_scores <- sapply(Endo2_stats$minus10_matches,
                                     function(x) PWMscoreStartingAt(pwm_s70, 
                                                                    subject(x),
                                                                    start(x)))

Endo2_stats$minus10_max_score <- sapply(Endo2_stats$minus10_scores,max)
Endo2_stats$minus10_max_entry <- sapply(Endo2_stats$minus10_scores,which.max)
Endo2_stats$minus10_scores <- sapply(Endo2_stats$minus10_matches,
                                     function(x) PWMscoreStartingAt(pwm_s70, 
                                                                    subject(x),
                                                                    start(x)))

for (i in seq(1,length(Endo2_stats$minus10_matches))) {
    if(Endo2_stats$minus10_max_score[i] == -Inf){
       Endo2_stats$minus10_max_score[i] <- 0
       Endo2_stats$minus10_start[i] <- -1
       Endo2_stats$minus10[i] <- "NNNNNN"
    }
    else{
      start_pos <- Endo2_stats$minus10_matches[i][[1]]@ranges@start[Endo2_stats$minus10_max_entry[[i]]]
      Endo2_stats$minus10_start[i] <- start_pos
      Endo2_stats$minus10[i] <- str_sub(Endo2_stats$variant[i],start_pos,start_pos+5)
    }
}

Endo2_stats$minus10_max <- NULL
Endo2_stats$minus10_max_entry <- NULL
```


Figure 1E cont. - Minus 35 Scoring

```{r warning=FALSE}
#This code uses the pwms from bTSSFinder (which come from literature) to find the best -35 site
A= c(0.0000,	0.0784,	0.0362,	0.4894,	0.3605,	0.4208)
C= c(0.1109,	0.0656,	0.0747,	0.2851,	0.3605,	0.0769)
G= c(0.1267,	0.0181,	0.6192,	0.1041,	0.0000,	0.2225)
T= c(0.7624,	0.8379,	0.2700,	0.1214,	0.2790,	0.2798)
pwm_s70_35 <- data.frame(A,C,G,T)
pwm_s70_35 <- t(pwm_s70_35)

Endo2_stats$minus35_matches <- sapply(as.character(Endo2_stats$variant),function(x) matchPWM(pwm_s70_35,str_sub(x,-150,-1))) #Normally (x,-150,-1)
Endo2_stats$minus35_scores <- sapply(Endo2_stats$minus35_matches,function(x) PWMscoreStartingAt(pwm_s70_35, subject(x),start(x)))
Endo2_stats$minus35_max_score <- sapply(Endo2_stats$minus35_scores,max)
Endo2_stats$minus35_max_entry <- sapply(Endo2_stats$minus35_scores,which.max)
Endo2_stats$minus35_scores <- sapply(Endo2_stats$minus35_matches,function(x) PWMscoreStartingAt(pwm_s70_35, subject(x),start(x)))

for (i in seq(1,length(Endo2_stats$minus35_matches))) {
    if(Endo2_stats$minus35_max_score[i] == -Inf){
       Endo2_stats$minus35_max_score[i] <- 0
       Endo2_stats$minus35_start[i] <- -1
       Endo2_stats$minus35[i] <- "NNNNNN"
    }
    else{
      start_pos <- Endo2_stats$minus35_matches[i][[1]]@ranges@start[Endo2_stats$minus35_max_entry[[i]]]
      Endo2_stats$minus35_start[i] <- start_pos
      Endo2_stats$minus35[i] <- str_sub(Endo2_stats$variant[i],start_pos,start_pos+5)
    }
}

Endo2_stats$minus35_max <- NULL
Endo2_stats$minus35_max_entry <- NULL

```


```{r}
a <- Endo2_stats %>% select(-tss_position, -strand) %>% na.omit() %>%
             mutate(minus35_start = -abs(150 - minus35_start)+30) %>%
             mutate(minus10_start = -abs(150 - minus10_start)+30) %>%
             select(minus10_max_score, minus35_max_score, classification) %>%
             filter(minus10_max_score > 0 & minus35_max_score >0) %>%
             ggplot(aes(x=classification, y=minus10_max_score)) +
             stat_boxplot(geom ='errorbar', width=.25) +
             geom_boxplot(aes(fill=classification), width=.5) +
             scale_fill_manual(values = c("#F6BA60","#2E6EB7", "grey")) +
             scale_x_discrete(expand = c(0, 2)) + 
             geom_signif(comparisons = list(c("Active", "Inactive"), c("Active", "Negative \n Controls"), 
                                            c("Inactive", "Negative \n Controls")), 
                         y_position=c(4.325, 4.4, 4.25), map_signif_level=TRUE) +
            labs(x="TSS Classification", y= "Minus10 Motif Score") +
            theme(axis.text.x = element_text(size = 15, angle = 45, hjust = 1),
        legend.position="none")

b <- Endo2_stats %>% select(-tss_position, -strand) %>% na.omit() %>%
             mutate(minus35_start = -abs(150 - minus35_start)+30) %>%
             mutate(minus10_start = -abs(150 - minus10_start)+30) %>%
             select(minus10_max_score, minus35_max_score, classification) %>%
             filter(minus10_max_score > 0 & minus35_max_score > 0) %>%
             ggplot(aes(x=classification, y=minus35_max_score)) + 
             stat_boxplot(geom ='errorbar', width=.25) +
             geom_boxplot(aes(fill=classification), width=.5) +
             scale_fill_manual(values = c("#F6BA60","#2E6EB7", "grey")) +
             labs(x="TSS Classification", y= "Minus35 Motif Score") +
             scale_x_discrete(expand = c(0, 2)) + 
             geom_signif(comparisons = list(c("Active", "Inactive"), 
                                            c("Active", "Negative \n Controls"), c("Inactive", "Negative \n Controls")), 
                         y_position=c(3.625, 3.7, 3.55), 
              map_signif_level=TRUE) + 
            theme(axis.text.x = element_text(size = 15, angle = 45, hjust=1),
        legend.position="none")

plot_grid(a,b, labels = c('A', 'B'))
ggsave('../../figs/fig1/tss_promoter_scores.pdf', units = 'in', width = 9.06, height = 6.32)

```











