---
title: "Figure2_V3"
author: "Guillaume Urtecho"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
setwd('~/Google Drive/Kosuri_Lab/ecoli_promoter_mpra/scripts/figs')
knitr::opts_knit$set(root.dir = '~/Google Drive/Kosuri_Lab/ecoli_promoter_mpra/scripts/figs')

options(stringsAsFactors = F)

library(dplyr)
library(ggplot2)
require(cowplot)
library(sqldf)
library(matrixStats)
library(reshape2)
library(reshape2)
library(Biostrings)
library(tidyverse)
library(scales)

options(scipen = 10000)

data_all <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                       header = T)

negatives <- filter(data_all, category == 'neg_control')

peak_data <- data_all %>% 
    filter(category == 'tile')
```

```{}
tfbs_regulondb <- read.table("../../ref/regulondb_tfbs_U00096.2.txt", 
                             skip = 30, fill = NA, sep = '\t') %>% 
    mutate(strand = ifelse(V6 == 'forward', '+', '-')) %>%
    select(tfbs = V2, tfbs_left = V4, tfbs_right = V5, strand) %>%
    distinct()
```

#How many active TSSs in promoter regions?
```{bash}

bedtools intersect -s -a ../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed -b ../../processed_data/endo_tss/lb/active_TSS_all.bed -wb | wc -l

#bedtools intersect -s -a ../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed -b ../../processed_data/endo_tss/lb/active_TSS_plus.bed -wo > ../../processed_data/endo_tss/lb/active_TSS_promoter_overlap_plus.bed

#bedtools intersect -s -a ../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed -b ../../processed_data/endo_tss/lb/active_TSS_minus.bed -wo > ../../processed_data/endo_tss/lb/active_TSS_promoter_overlap_minus.bed

```

How many inactive TSSs in promoter regions?
```{bash}

bedtools intersect -s -a ../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed -b ../../processed_data/endo_tss/lb/inactive_TSS_all.bed -wb | wc -l

```
How many negative controls in promoter regions?
```{bash}

bedtools intersect -s -a ../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed -b ../../processed_data/endo_tss/lb/negative_TSS_plus.bed -wb #| wc -l

```


2C - Fragment Activity at TSSs

```{r}
is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))

positions <- seq(10, 1150, 10) - 620 

plus_tss <- read.table("../../processed_data/frag/lb/plus_tss_counts.txt", 
                       skip = 3, fill = TRUE, sep = '\t', header = F)
plus_tss[is.nan(plus_tss)] <- 0
names(plus_tss) <- positions

plus_tss_active <-plus_tss[c(1:1009), ]
plus_tss_inactive <-plus_tss[c(1010:8231), ]
plus_tss_negative <- plus_tss[c(8232:8705), ]

minus_tss <- read.table("../../processed_data/frag/lb/minus_tss_counts.txt", 
                        skip =3, sep = '\t', header = F)
minus_tss[is.nan(minus_tss)] <- 0

names(minus_tss) <- positions
minus_tss_active <-minus_tss[c(1:1002), ]
minus_tss_inactive <-minus_tss[c(1003:8476), ]

tss_active <- rbind(plus_tss_active, minus_tss_active) %>%
    melt() %>%
    mutate(variable = as.numeric(as.character(variable)),
           type = 'Active TSS')
                  
names(tss_active) <- c('position', 'coverage', 'Type')

tss_inactive <- rbind(plus_tss_inactive, minus_tss_inactive) %>%
    melt(by = 'type') %>%
    mutate(variable = as.numeric(as.character(variable)),
           type = 'Inactive TSS')
names(tss_inactive) <- c('position', 'coverage', 'Type')

tss_negative <- plus_tss_negative %>%
        melt() %>%
        mutate(variable = as.numeric(as.character(variable)),
               type = 'Negative TSS')

names(tss_negative) <- c('position', 'coverage', 'Type')
```

```{r}
rbind(tss_active, tss_inactive, tss_negative) %>% 
    ggplot(., aes(x = position, y = coverage)) +
    geom_smooth(method ='auto', aes(fill = Type, color = Type)) +
    geom_vline(xintercept = c(1), linetype = 'dashed') +
    scale_color_manual(breaks = c('Active TSS', 'Inactive TSS', "Negative TSS"),
                     values = c("#2E6EB7", "#F6BA60", "grey")) +
    scale_fill_manual(breaks = c('Active TSS', 'Inactive TSS', "Negative TSS"),
                     values = c("#2E6EB7", "#F6BA60", "grey")) +
    labs(x ='Position Relative to TSS',
         y = 'Mean Promoter Activity\n(per Base)') +
    theme(axis.title.x = element_text(size = 20),
        axis.text.y = element_text(size = 20), 
        axis.text.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20), 
        legend.text = element_text(size = 20),
        legend.title = element_blank())

ggsave('../../figs/fig2/tss_frag_activity.pdf', units = 'in',
       height = 4.5, width = 8)
```



Figure3B, MraZ Promoter

```{r}
threshold <- 1
peak_data %>% 
    filter(grepl('89077_89780', name)) %>% 
    ggplot(aes(tile_start + peak_start, expn_med_fitted_scaled)) + 
    geom_point(aes(color = active)) +
    scale_y_continuous(breaks = seq(0.5, 3, 0.5)) +
    geom_hline(yintercept = threshold, linetype = 'dashed') +
    geom_line(linetype = 'dashed') +
    scale_color_manual(values = c('red', 'black')) +
    labs(x = 'Tile Start Position', y = 'Expression', color = '', title = "mraZ Promoter") +
    annotate("text", label = "Threshold", x = 89125, y = threshold+.1) +
    theme(axis.text = element_text(size = 20),
          axis.title = element_text(size = 20),
          title = element_text(size = 24),
          legend.text = element_text(size = 20))

ggsave("../../figs/fig2/MraZPromoter.pdf", units = 'in',
       height = 5, width = 7)
```

Subset active promoters and print for merging
```{r}
peak_data %>% 
    filter(expn_med_fitted_scaled > threshold) %>% 
    nrow()  #15013 sequences with more active expression

peak_data %>% 
    filter(expn_med_fitted_scaled > threshold) %>%
    mutate(genome = "U00096.2",
           left = as.integer(tile_start+peak_start),
           right = tile_end+peak_start) %>% arrange(left) %>%
    select(genome, left, right, name, expn_med_fitted_scaled, strand) %>%
    write.table(., "../../processed_data/peak_tile/active_tiles.bed", 
                row.names = F, quote = F, sep = '\t', col.names = FALSE)
```


Identify overlapping active regions

```{bash}
bedtools merge -i ../../processed_data/peak_tile/active_tiles.bed -d -100 -s -c 5,6 -o median,distinct > ../../processed_data/peak_tile/merged_active_tiles.bed
```

```{r}
read.table("../../processed_data/peak_tile/merged_active_tiles.bed") %>%
    mutate(name = paste(V2, "_", V3, "_", V5, sep = "")) %>%
    select(V1, V2, V3, name, V4, V5) %>%
    write.table(., "../../processed_data/peak_tile/merged_active_tiles.bed",
                  col.names = FALSE,
                  sep = '\t',
                  row.names = F,
                  quote = F)
```

Identify active tiles that overlap peaks
```{bash}
bedtools intersect -c -a ../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed -b ../../processed_data/peak_tile/merged_active_tiles.bed -s > ../../processed_data/peak_tile/collapsed_per_peak.bed
                                    
```

3C - number of promoters per peak
```{r}
collapsed_per_Peak <- read.table("../../processed_data/peak_tile/collapsed_per_peak.bed",
                                 col.names = c("chrom", 'start', 'end', 'name', 'peak_score',
                                               'strand', 'score1', 'score2', 'num_sub_peaks')) %>% 
    separate(name, into = c('peak_start', 'peak_end', 'strand'), sep = '_', remove = F, convert = T) %>% 
    # only consider peaks with mapped tiles, eliminates 123 peaks
    semi_join(peak_data, by = c('peak_start', 'peak_end'))

collapsed_per_Peak %>% 
    filter(num_sub_peaks < 10) %>% 
    ggplot(aes(num_sub_peaks)) +
        geom_bar(stat = "count", width = .8, fill = '#528ECB', color = 'black') +
        geom_text(stat = 'count', aes(label = ..count..), vjust = -0.2, size = 10) +
        scale_y_log10() + annotation_logticks(sides = 'l') + 
        labs(x = 'Number of promoters per region', y = "Count") + 
        scale_x_continuous(breaks = pretty_breaks(n = 7)) +
        theme(axis.text = element_text(size = 20),
              axis.title = element_text(size = 20),
              legend.text = element_text(size = 20))

ggsave("../../figs/fig2/promoters_per_peak.pdf", units = 'in',
       height = 4, width = 6)

#How many promoters do we identify?

sum(collapsed_per_Peak$num_sub_peaks) #2,228

```


3D) Identify minimal promoter unit, 
```{r}
minimal_promoters <- read.table("../../processed_data/peak_tile/merged_active_tiles.bed", header = F,
                                col.names = c("genome", 'left', 'right', 'name', 'score', 'strand')) %>% 
                                mutate(left_minimal = right - 150,
                                right_minimal = left + 150,
                                minimal_length = right_minimal - left_minimal,
                                name = paste(left_minimal, "_", right_minimal, "_", strand, sep = '')) %>%
                                filter(right_minimal - left_minimal > 0)

minimal_promoters %>% 
    select(genome, left_minimal, right_minimal, name,score, strand) %>%
    write.table(., "../../processed_data/peak_tile/minimal_promoters.bed", 
                row.names = F, quote = F, col.names = FALSE, sep = '\t')

# sizes of minimal promoters
a <- ggplot(filter(minimal_promoters, minimal_length > 0), aes(minimal_length)) + 
    geom_histogram(color = 'black',
                   binwidth = 10,
                   fill = '#528ECB',
                   color = 'black') +
    scale_x_continuous(breaks = pretty_breaks(n=8)) + 
    labs(x = 'Length of minimum promoter',
         y = 'Count',
         title = 'Sizes of minimum promoter regions') +
    geom_point(shape = 25, fill = 'firebrick3', color = 'navyblue', 
               aes(x = 40, y = 300), size = 3) +
    geom_point(shape = 25, fill = 'navyblue', color = 'navyblue', 
               aes(x = 150, y = 450), size = 3) +
    theme(axis.text = element_text(size = 20),
          axis.text.x = element_text(size = 20, angle = 45, hjust = 1),
          axis.title = element_text(size = 20),
          legend.text = element_text(size = 20))

b <- minimal_promoters %>% 
    filter(minimal_length > 35 & minimal_length < 45) %>%
    ggplot(aes(score)) + 
    geom_density(color = 'firebrick3', alpha = 0.4) +
    scale_x_log10(limits = c(NA, 30)) + 
    labs(y = 'Density', x = '')

c <- minimal_promoters %>% 
    filter(minimal_length == 150) %>%
    ggplot(aes(score)) + 
    geom_density(color = 'navyblue', alpha = 0.4) +
    scale_x_log10(limit = c(NA,30)) + 
    labs(y = 'Density', x = 'Minimal Promoter Expression')

dens <- plot_grid(b,c, ncol = 1)
plot_grid(a,dens, ncol = 2, rel_heights = c(1, .8))


ggsave("../../figs/fig2/minimal_promoter_size.pdf", units = 'in', 
       width = 7, height = 5)
```




#So how many total promoters? add 2252 minimal promoters
```{bash}

#Number only in tss library
bedtools intersect -v -a ../../processed_data/endo_tss/lb/active_TSS_all.bed -b ../../processed_data/peak_tile/merged_active_tiles.bed -s | wc -l

```

