---
title: "Classification Models"
output: html_notebook
---

```{r}
load_pkgs <- function(pkgs){
    new_pkgs <- pkgs[!(pkgs %in% installed.packages()[, 'Package'])]
    if(length(new_pkgs)) install.packages(new_pkgs)
    for(pkg in pkgs){
        suppressWarnings(suppressMessages(library(pkg, character.only = T)))
    }
}

pkgs <- c('dplyr', 'ggplot2', 'tidyr', 'cowplot', 'ggsignif', 'stringr', 'pracma',
          'PRROC', 'wesanderson')
load_pkgs(pkgs)

library(Biostrings)

options(stringsAsFactors = F)
```

Calculate PWM scores for TSS library only.

```{r}
data <- read.table('../../processed_data/endo_tss/lb/tss_expression_pwm_info.txt',
                  header = T) %>% 
    filter(expn_med_fitted_scaled < 0.75 | expn_med_fitted_scaled >= 1.25) %>% 
    mutate(label = ifelse(expn_med_fitted_scaled <= 0.75, 0, 1))
data$label <- factor(data$label)

calc_gc <- function(seq) {
    seq_length <- nchar(seq)
    noG <- gsub('G', '', seq)
    noGC <- gsub('C', '', noG)
    numGC <- seq_length - nchar(noGC)
    gc_content <- numGC / seq_length
    return(gc_content)
}

data$gc_content <- calc_gc(data$variant)

train <- filter(data, dataset == 'train')
test <- filter(data, dataset == 'test')
# simple linear model based on max -35 and -10 score
minus35and10_logit_fit <- glm(label ~ minus35_max_score + minus10_max_score +
                                  pwm_paired_max + gc_content,
                              data = train, family = 'binomial')

summary(minus35and10_logit_fit)
```


```{r}
test$predicted <- predict(minus35and10_logit_fit, test, type = 'response')
logit_curve <- pr.curve(scores.class0 = test$predicted[test$label == 1],
                        scores.class1 = test$predicted[test$label == 0], curve = T)
plot(logit_curve)
```

Create training sets for 75% of the genome and test on remaining 25%. Combine TSS,
scramble and peak libraries. Points near our threshold of 1.0 could be noisy, 
so let's only consider negatives as < 0.75 and positives as > 1.25

Training: 12,918 positives and 40,408 negatives = 53,326
Test: 4,414 positives and 14,153 negatives = 18,567 

```{r}
calculate_precision <- function(df) {
    # how many predicted positives are true positives
    num_false_positive <- filter(df, predicted == 1, observed == 0) %>% nrow()
    num_true_positive <- filter(df, predicted == 1, observed == 1) %>% nrow()
    precision <- num_true_positive / (num_true_positive + num_false_positive)
    return(precision)
}

calculate_recall <- function(df) {
    # how many true positives are predicted positive
    num_true_positive <- filter(df, predicted == 1, observed == 1) %>% nrow()
    num_false_negative <- filter(df, predicted == 0, observed == 1) %>% nrow()
    recall <- num_true_positive / (num_true_positive + num_false_negative)
    return(recall)
}
```

```{r}
logistic <- read.table('../../processed_data/combined/20190613_kmer_logistic_predictions.txt',
                       col.names = c('predicted', 'observed'))
logistic_curve <- pr.curve(scores.class0 = logistic$predicted[logistic$observed == 1],
                           scores.class1 = logistic$predicted[logistic$observed == 0],
                           curve = T)
plot(logistic_curve)
```


```{r}
pls <- read.table('../../processed_data/combined/20190613_kmer_pls_classification_predictions.txt',
                       col.names = c('predicted', 'observed'))
pls_curve <- pr.curve(scores.class0 = pls$predicted[pls$observed == 1],
                           scores.class1 = pls$predicted[pls$observed == 0],
                           curve = T)
plot(pls_curve)
```

```{r}
rf <- read.table('../../processed_data/combined/20190620_rf_tss_scramble_peak_classification_predictions.txt',
                              col.names = c('predicted', 'observed'))
rf_curve <- pr.curve(scores.class0 = rf$predicted[rf$observed == 1],
                     scores.class1 = rf$predicted[rf$observed == 0],
                     curve = T)
plot(rf_curve)
```


```{r}
rf_onehot <- read.table('../../processed_data/combined/20190611_rf_onehot_tss_scramble_peak_classification_predictions.txt',
                              col.names = c('predicted', 'observed')) %>% 
    mutate(predicted = as.numeric(predicted),
           observed = as.numeric(observed)) %>% 
    filter(!is.na(predicted), !is.na(observed))
rf_onehot_curve <- pr.curve(scores.class0 = rf_onehot$predicted[rf_onehot$observed == 1],
                            scores.class1 = rf_onehot$predicted[rf_onehot$observed == 0],
                            curve = T)
plot(rf_onehot_curve)
```

```{r}
mlp <- read.table('../../processed_data/combined/20190611_mlp_tss_scramble_peak_predictions_classification.txt',
                  col.names = c('predicted', 'observed'))
mlp_curve <- pr.curve(scores.class0 = mlp$predicted[mlp$observed == 1],
                      scores.class1 = mlp$predicted[mlp$observed == 0],
                      curve = T)
plot(mlp_curve)
```


```{r}
mlp_3to6 <- read.table('../../processed_data/combined/20190612_mlp_3to6mer_tss_scramble_peak_predictions_classification.txt',
                  col.names = c('predicted', 'observed'))
mlp_3to6_curve <- pr.curve(scores.class0 = mlp_3to6$predicted[mlp_3to6$observed == 1],
                           scores.class1 = mlp_3to6$predicted[mlp_3to6$observed == 0],
                           curve = T)
plot(mlp_3to6_curve)
```

"The no-skill line is defined by the total number of positive cases divide by the total number of positive and negative cases."

```{r}
nn <- read.table('../../processed_data/combined/20190610_tss_scramble_peak_classification_hyperparam_tuned_predictions.txt',
                 col.names = c('predicted', 'observed'))
nn_curve <- pr.curve(scores.class0 = nn$predicted[nn$observed == 1],
                     scores.class1 = nn$predicted[nn$observed == 0], curve = T)
plot(nn_curve)

```


```{r}
# gkMSVM for TSS only
gkmsvm <- read.table('../../processed_data/endo_tss/lb/model_files/gkmsvm_10mer_8ungapped_kernel_test_results',
                     col.names = c('name', 'predicted'))
tss_test <- read.table('../../processed_data/endo_tss/lb/model_files/tss_expression_model_format_test_genome_split.txt',
                   header = F, sep = '\t', col.names = c('sequence', 'expression'))
gkmsvm$observed <- tss_test$expression

gkmsvm_pr_curve <- pr.curve(scores.class0 = filter(gkmsvm, observed >= 1)$predicted,
                     scores.class1 = filter(gkmsvm, observed < 1)$predicted,
                     curve = T)
plot(gkmsvm_pr_curve)
```

```{r}
tp_rate <- 4414/18567

pdf('../../figs/fig6/model_comparison_classification.pdf')
par(pty = 's')
plot(logit_curve, add=F, color = 'red', auc.main = F, main = '')
plot(logistic_curve, add = T, color = 'aquamarine')
plot(pls_curve, add = T, color = 'black')
plot(rf_curve, add=T, color = 'blue')
plot(rf_onehot_curve, add = T, color = 'darkgreen')
plot(mlp_curve, add = T, color = 'lightblue')
plot(mlp_3to6_curve, add = T, color = 'purple')
plot(nn_curve, add = T, color = 'gold')
plot(gkmsvm_pr_curve, add = T, color = 'grey', legend = T)
abline(h = tp_rate, lty = 'dashed')
dev.off()
```

```{r}
combined_pct_pos <- length(which(mlp$observed == 1)) / nrow(mlp)
tss_pct_pos <- length(which(gkmsvm$observed > 1)) / nrow(gkmsvm)
print(c(combined_pct_pos, tss_pct_pos))
```

