---
title: "Endo_supplement"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = '~/Google Drive/Kosuri_Lab/Endo_Paper/scripts/')
#setwd("C:/Users/george/Google Drive/Kosuri_Lab/Endo_Paper/scripts")
#setwd("~/../Google Drive/Kosuri_Lab/ecoli_promoter_mpra (1)/scripts/")
#setwd("~/Google Drive/Kosuri_Lab/Endo_Paper/scripts/")
setwd('~/Google Drive/Kosuri_Lab/ecoli_promoter_mpra/scripts/figs')
knitr::opts_knit$set(root.dir = '~/Google Drive/Kosuri_Lab/ecoli_promoter_mpra/scripts/figs')
library("ggplot2")
library("dplyr")
library("tidyr")
require("cowplot")
library("bbplot")
library(psych)
library(reshape2)
library(matrixStats)
library(sqldf)
library(ggsignif)
#library('png')

options(stringsAsFactors = F)
options(scipen = 10000)

```


For each of these 18,000 TSS promoters and controls we generated an average of 50 reporter constructs in which each construct expressed a uniquely barcoded GFP (Figure S1)

Figure S1
```{r}
Endo2 <- read.table("../../processed_data/endo_tss/lb/rLP5_Endo2_lb_expression_formatted_std.txt", header = T, fill  = T)

#Figure S1A - Distribution of the number of barcodes per TSS
median_barcodes <- median(Endo2$num_barcodes_integrated)

d <- ggplot(Endo2, aes(x=num_barcodes_integrated)) +
      geom_histogram(color = 'black') +
      scale_x_log10() +
      labs(x = "Number of barcodes",
           y = 'Counts') +
      annotation_logticks(sides = 'b') +
      geom_vline(xintercept = median_barcodes, color = 'red', linetype = 'dashed') +
      annotate('text', x = 10, y = 1500, label = paste("Median =", median_barcodes, "barcodes"), size = 8) +
    theme(axis.text.x = element_text(size = 24), 
        axis.text.y = element_text(size = 24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))

#Figure S2B - Pairwise comparisons on different landing pad locations
fLP3 <- read.table("../../processed_data/endo_tss/alt_landing_pads/fLP3/fLP3_Endo2_lb_expression_formatted_std.txt", header = T) %>% 
      #select(name, fLP3=expn_med_fitted_scaled) %>%
      select(name, fLP3=RNA_exp_sum_ave) %>%
      mutate(fLP3 = log10(as.numeric(fLP3)))

rLP6 <- read.table("../../processed_data/endo_tss/alt_landing_pads/rLP6/rLP6_Endo2_lb_expression_formatted_std.txt", header = T) %>%
          #select(name, rLP6=expn_med_fitted_scaled) %>%
          select(name, rLP6=RNA_exp_sum_ave) %>%
          mutate(rLP6 = log10(as.numeric(rLP6)))

compare_locations <- inner_join(Endo2, fLP3, by = 'name') %>% 
    inner_join(., rLP6, by = 'name') %>%
    select('Right Midreplichore'=fLP3,'Terminus'=RNA_exp_sum_ave, 'Left Midreplichore' = rLP6, 'name') %>% 
    mutate(Terminus = log10(Terminus))

neg <- subset(compare_locations, grepl("neg_control", compare_locations$name))
pos <- subset(compare_locations, grepl("pos_control", compare_locations$name))

#Right vs Terminus
cor <- cor(compare_locations$`Right Midreplichore`, compare_locations$Terminus)

a <- ggplot(compare_locations, aes(`Right Midreplichore`, Terminus)) + 
  geom_point(alpha = .1) +
  annotate("text", label = paste("italic(r)==", signif(cor, 3)), parse = TRUE,
           x = 0, y = 1.75, size = 10) +
  xlab('') +
  ylab('Log10(terminus expression)') +
  geom_point(data = neg,aes(`Right Midreplichore`, Terminus), color = "firebrick1") +
  geom_point(data = pos,aes(`Right Midreplichore`, Terminus), color = "dodgerblue1") +
  annotate("text", label = "Negative Controls", x = 1.5, y = 0, color = 'firebrick1', size = 8) +
  theme(axis.text.x = element_text(size = 24), 
        axis.text.y = element_text(size = 24),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20))

#Right vs left 
cor <- cor(compare_locations$`Right Midreplichore`, compare_locations$`Left Midreplichore`)

b <- ggplot(compare_locations, aes(`Right Midreplichore`, `Left Midreplichore`)) + 
  geom_point(alpha = .1) +
  annotate("text", label = paste("italic(r)==", signif(cor, 3)), parse = TRUE,
           x = 0, y = 1.5, size = 10) +
  xlab('Log10(right midreplichore expression)') +
  ylab('Log10(left midreplichore expression)') +
  geom_point(data = neg,aes(`Right Midreplichore`, `Left Midreplichore`), color = "firebrick1") +
  geom_point(data = pos,aes(`Right Midreplichore`, `Left Midreplichore`), color = "dodgerblue1") +
  annotate("text", label = "Negative Controls", x = 1.5, y = .1, color = 'firebrick1', size = 8) +
  theme(axis.text.x = element_text(size = 24), 
        axis.text.y = element_text(size = 24),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20))

#Left vs Terminus
cor <- cor(compare_locations$`Left Midreplichore`, compare_locations$Terminus)

c <- ggplot(compare_locations, aes(Terminus, `Left Midreplichore`)) + 
  geom_point(alpha = .1) +
  annotate("text", label = paste("italic(r)==", signif(cor, 3)), parse = TRUE,
           x = 0, y = 1.5, size = 10) +
  xlab('Log10(terminus expression)') +
  ylab('') +
  geom_point(data = neg,aes(Terminus, `Left Midreplichore`), color = "firebrick1") +
  geom_point(data = pos,aes(Terminus, `Left Midreplichore`), color = "dodgerblue1") +
  annotate("text", label = "Negative Controls", x = 1.5, y = .1, color = 'firebrick1', size = 8) +
  theme(axis.text.x = element_text(size = 24), 
        axis.text.y = element_text(size = 24),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20))


e <-plot_grid(a,NULL,b,c)

plot_grid(d,e, ncol = 1, labels = 'AUTO',rel_widths = c(1,3), rel_heights = c(1,3), label_size = 18)

ggsave('../../figs/supp_figs/S1_TSS.stats.png', units = 'in', width = 12.29, height = 16)


```
```{r}
tss <- read.table('../../processed_data/endo_tss/lb/rLP5_Endo2_lb_expression_formatted.txt', 
                  header = T)

scramble <- read.table('../../processed_data/endo_scramble/endo_scramble_expression_formatted.txt', 
                       header = T)
# graph unscrambled
wildtype <- inner_join(filter(tss, category == 'tss') %>% 
                           select(name, tss_name, expn_med1 = expn_med),
                       filter(scramble, category == 'unscrambled') %>% 
                           select(name, tss_name, expn_med2 = expn_med),
                       by = 'tss_name')


r2 <- summary(lm(expn_med1 ~ expn_med2, wildtype))$r.squared
ggplot(wildtype, aes(expn_med1, expn_med2)) + 
    geom_point() + 
    annotation_logticks(sides='bl') + scale_x_log10() + scale_y_log10() +
    geom_smooth(method = 'lm') +
    labs(x = 'Barcode Replicate #1', y = 'Barcode replicate #2',
         title = '') +
    annotate('text', x=50, y=0.1, parse=T, label=paste('R^2==', signif(r2, 3)), size = 8) +
            theme(axis.text.x = element_text(size = 20), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20))


```


Figure S1B - COmparing alternative LP locations
```{} 
fLP3 <- read.table("../../processed_data/endo_tss/alt_landing_pads/fLP3/fLP3_Endo2_lb_expression_formatted_std.txt", header = T) %>% 
      select(name, fLP3=expn_med_fitted_scaled) %>%
      mutate(fLP3 = log10(as.numeric(fLP3)))



rLP6 <- read.table("../../processed_data/endo_tss/alt_landing_pads/rLP6/rLP6_Endo2_lb_expression_formatted_std.txt", header = T) %>%
          select(name, rLP6=expn_med_fitted_scaled) %>%
          mutate(rLP6 = log10(as.numeric(rLP6)))

compare_locations <- inner_join(Endo2, fLP3, by = 'name') %>% 
    inner_join(., rLP6, by = 'name') %>%
    select('Right Midreplichore'=fLP3,'Terminus'=expn_med_fitted_scaled, 'Left Midreplichore' = rLP6) %>% 
    mutate(Terminus = log10(Terminus))


pairs.panels(compare_locations, 
             method = "pearson", 
             hist.col = "#00AFBB",
             lm = T,
             breaks = 30,
             rug = F,
             density = FALSE,
             cex.axis = 3,
             stars = T,
             cex.labels = 3, font.labels =3)


ggsave('../../figs/supp_figs/S1B_ReplicatesBetweenLocation.png', units = 'in', width = 7.29, height = 4.51)
```




Figure S2 - Genome fragment screen for promoters
```{r}
frag_expression <- read.table("../../processed_data/frag/lb/U00096.2_frag-rLP5_LB_expression.txt", 
           header = T) %>% mutate(frag_length = end - start)


cor <- cor(frag_expression$RNA_exp_1, frag_expression$RNA_exp_2)

a <- ggplot(frag_expression, aes(RNA_exp_1, RNA_exp_2)) + 
  geom_point(alpha = .1) +
  scale_x_log10() + scale_y_log10() + annotation_logticks(sides= 'bl') +
  annotate("text", label=paste("italic(r)==",signif(cor, 3)),parse=TRUE, x=.5, y=15, size = 8) +
  xlab('Biological Replicate #1') + ylab('Biological Replicate #2') +
  theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))
median_frag_length <- median(frag_expression$frag_length)
sum_fragments <- sum(frag_expression$frag_length)

b <- ggplot(frag_expression, aes(frag_length)) + 
      geom_histogram(color = 'black') +
      labs(x = "Length of Fragment (bp)",
            y = 'Counts') +
      geom_vline(xintercept = median_frag_length, color = 'red', linetype = 'dashed') +
      annotate('text', x = 400, y = 50000, label = paste("Median =", median_frag_length, "bp"), size = 6.5) +
  annotate('text', x = 400, y = 45000, label = paste(signif(sum_fragments/(4639675*2), 3), "x", " coverage", sep = ''), size = 6.5) +
  theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20)) 


# read in barcode count files
filelist = list.files(path = "../../processed_data/frag/lb",
                      pattern = '*us_rep*',
                      full.names = T)

# should be named so files are in following order, number is biological replicate

sample_names = c('Minus_1', 'Minus_2', 'Plus_1', 'Plus_2')

for(i in seq(1:length(filelist))) {
    sample_name <- sample_names[i]
    x <- read.table(filelist[i], col.names=c('position', sample_name), header = F, skip = 1)
    x[[sample_name]] <- 1000000*x[[sample_name]]/sum(x[[sample_name]])  #Normalizes by RPM
    assign(sample_name, x)  
}

count_list <- list(Minus_1, Minus_2, Plus_1, Plus_2)

set.seed(123)
nucleotide_exp <- count_list %>% purrr::reduce(full_join, by = "position") %>% sample_n(50000, replace = F) %>% 
  melt(id = 'position') %>% 
  separate(variable, by = '_', into = c('strand', 'replicate')) %>% 
  spread(key = 'replicate', value = 'value')

cor <- cor(nucleotide_exp$`1`, nucleotide_exp$`2`)    

c <- ggplot(nucleotide_exp, aes(`1`, `2`)) + 
  geom_point(alpha = .1) +
  scale_x_log10() + scale_y_log10() + annotation_logticks(sides= 'bl') +
  annotate("text", label=paste("italic(r)==",signif(cor, 3)),parse=TRUE, x=.1, y=5, size = 8) +
  xlab('Biological Replicate #1') + ylab('Biological Replicate #2') +
  theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))


#Figure S2D

overlap_plus <- read.table("../../processed_data/endo_tss/lb/active_TSS_promoter_overlap_plus.bed") %>% select(name=V12)
overlap_combined <- read.table("../../processed_data/endo_tss/lb/active_TSS_promoter_overlap_minus.bed") %>% select(name=V12) %>% 
                      rbind(., overlap_plus) %>% distinct() %>% mutate(capture = 'Captured')
d<- read.table('../../processed_data/endo_tss/lb/rLP5_Endo2_lb_expression_formatted_std.txt', header = T) %>% filter(active == 'active') %>%
              left_join(., overlap_combined, by ='name') %>% replace(., is.na(.), 'Missing') %>% filter(category == 'tss') %>%
        ggplot(aes(x=capture, y=expn_med_fitted_scaled)) +
        geom_boxplot(fill = 'gray60', alpha = .6) +
        scale_y_log10() +
        annotation_logticks(sides = 'l') +
        geom_hline(yintercept = 1, linetype = 'dashed', color = 'red') + 
        labs(x='', y = 'TSS promoter activity') +
          theme(axis.text.x = element_text(size = 20), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20))

#how many captured vs missing?
read.table('../../processed_data/endo_tss/lb/rLP5_Endo2_lb_expression_formatted_std.txt', header = T) %>% filter(active == 'active') %>%
              left_join(., overlap_combined, by ='name') %>% replace(., is.na(.), 'Missing') %>% filter(category == 'tss') %>%
              select(capture) %>%
              group_by(capture) %>%
              mutate(num_tss = n()) %>% ungroup() %>% distinct()
  


##Figure S2E 
peak_lengths <- read.table('../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed',
           col.names = c('genome', 'peak_start', 'peak_end', 'name', 'score', 'strand', 'peakscore', 'peak_max')) %>% select(name, peak_start, peak_end) %>% mutate(peak_length = peak_end-peak_start)

med_peak_length <- median(peak_lengths$peak_length)

e <- ggplot(peak_lengths, aes(x=peak_length)) +
    geom_histogram(color = 'black') +
      labs(x = "Length of Peak (bp)",
            y = 'Counts') +
    scale_x_log10() +
    annotation_logticks(sides = 'b') +
      geom_vline(xintercept = med_peak_length, color = 'red', linetype = 'dashed') +
      annotate('text', x = 500, y = 300, label = paste("Median =", med_peak_length, "bp"), size = 8) +
  theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20)) 


#Figure S2F 
peak_data <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt', header = T)

neg <- subset(peak_data, grepl("neg_control", peak_data$name))
threshold <- 1


cor <- cor(peak_data$RNA_exp_sum_1, peak_data$RNA_exp_sum_2)

f <- ggplot(peak_data, aes(RNA_exp_sum_1, RNA_exp_sum_2)) + 
  geom_point(alpha = .1) +
  scale_x_log10() + scale_y_log10() + annotation_logticks(sides= 'bl') +
  annotate("text", label=paste("italic(r)==",signif(cor, 3)),parse=TRUE, x=.5, y=15, size = 8) +
  xlab('Biological Replicate #1') + ylab('Biological Replicate #2') +
  geom_point(data=neg,aes(RNA_exp_sum_1, RNA_exp_sum_2), color = "firebrick1") +
  annotate("text", label = "Negative\nControls", x = 20, y = .5, color = 'firebrick1', size = 8) +
  theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))


#Figure S2G

median_barcodes <- median(peak_data$num_barcodes_integrated)

g <- ggplot(peak_data, aes(x=num_barcodes_integrated)) +
      geom_histogram(color = 'black') +
      labs(x = "Number of barcodes",
            y = 'Counts') +
      geom_vline(xintercept = median_barcodes, color = 'red', linetype = 'dashed') +
      annotate('text', x = 50, y = 5000, 
               label = paste("Median =", median_barcodes, "barcodes"), size = 7) +
       theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))





#Figure S2H

h <- read.table("../../processed_data/peak_tile/collapsed_per_peak.bed") %>%
    mutate(length = V3-V2) %>% select(length, V9) %>%
    ggplot(aes(as.character(V9),length, fill = as.character(V9))) +
    geom_boxplot() +
    viridis::scale_fill_viridis(discrete = T) +
    labs(x = 'Number of promoters\nin region',
         y = 'Promoter region size (bp)',
         fill = 'Number of promoters\nin region') + 
    geom_hline(yintercept = 150, linetype = 'dashed', color = 'red') +
    annotate('text', x = '6', y = 220, label = '150 bp', color = 'red', size = 6, hjust = .75) +
    theme(legend.position = 'none',
        axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20))


#Figure S2I

i <- read.table("../../processed_data/peak_tile/collapsed_per_peak.bed") %>%
    mutate(length = V3-V2, pbActivity = V5/length) %>% select(length, V9, pbActivity) %>%
    ggplot(aes(as.character(V9), pbActivity,  fill = as.character(V9))) +
    geom_boxplot() +
    viridis::scale_fill_viridis(discrete = T) +
    scale_y_log10() +
    labs(x = 'Number of promoters\nin region',
         y = 'Per base promoter activity',
         fill = 'Number of Promoters\nin region') +
    geom_hline(yintercept = 1.1, linetype = 'dashed', color = 'red') +
    annotate('text', x = '6', y = 1.3, label = 'Threshold', color = 'red', size = 6, hjust = .75) +
    theme(legend.position = 'none',
        axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.title = element_text(size = 20))

plot_grid(a,b,c,d,e,f,g,h,i, ncol = 3, labels = 'AUTO', label_size = 20)

ggsave('../../figs/supp_figs/S2_GenomicPromoters.png', units = 'in', width = 16, height = 16)


```

Figure S4 - QC of frag in M9
```{r}
frag_expression <- read.table("../../processed_data/frag/m9/U00096.2_frag-rLP5_M9_expression.txt", 
           header = T) %>% mutate(frag_length = end - start)

#S4A
cor <- cor(frag_expression$RNA_exp_1, frag_expression$RNA_exp_2)

a <- ggplot(frag_expression, aes(RNA_exp_1, RNA_exp_2)) + 
  geom_point(alpha = .1) +
  scale_x_log10() + scale_y_log10() + annotation_logticks(sides= 'bl') +
  annotate("text", label=paste("italic(r)==",signif(cor, 3)),parse=TRUE, x=.5, y=15, size = 8) +
  xlab('Biological Replicate #1') + ylab('Biological Replicate #2') +
  theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))

median_frag_length <- median(frag_expression$frag_length)
sum_fragments <- sum(frag_expression$frag_length)

#S4B
b <- ggplot(frag_expression, aes(frag_length)) + 
      geom_histogram(color = 'black') +
      labs(x = "Length of Fragment (bp)",
           y = 'Counts') +
      geom_vline(xintercept = median_frag_length, color = 'red', linetype = 'dashed') +
      annotate('text', x = 400, y = 50000,
               label = paste("Median =", median_frag_length, "bp"), size = 8) +
      annotate('text', x = 400, y = 45000,
               label = paste(signif(sum_fragments/(4639675*2), 3), "x", " coverage", sep = ''), size = 8) +
      theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20)) 



##Figure S4C 
peak_lengths <- read.table('../../processed_data/frag/m9/U00096.2_M9_plus_minus_called_peaks_threshold1.1_merge40_min60.bed',
           col.names = c('genome', 'peak_start', 'peak_end', 'name', 'score', 'strand', 'peakscore', 'peak_max')) %>% select(name, peak_start, peak_end) %>% mutate(peak_length = peak_end-peak_start)

med_peak_length <- median(peak_lengths$peak_length)

c <- ggplot(peak_lengths, aes(x=peak_length)) +
    geom_histogram(color = 'black') +
      labs(x = "Length of Peak (bp)",
            y = 'Counts') +
    scale_x_log10() +
    annotation_logticks(sides = 'b') +
      geom_vline(xintercept = med_peak_length, color = 'red', linetype = 'dashed') +
      annotate('text', x = 650, y = 400, label = paste("Median =", med_peak_length, "bp"), size = 8) +
  theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20)) 

#S4D, unique peaks are often lower expressed than shared peaks

# Identify peaks that are unique to each condition

M9_peaks <- read.table("../../processed_data/frag/m9/U00096.2_M9_plus_minus_called_peaks_threshold1.1_merge40_min60.bed",
                col.names = c('genome', 'peak_start', 'peak_end', 'name', 'score', 'strand', 'peakscore', 'peak_max')) %>%
                select(name, peak_start, peak_end, score) %>% mutate(peak_length = peak_end-peak_start, condition = 'M9')

LB_peaks <- read.table("../../processed_data/frag/lb/U00096.2_plus_minus_called_peaks_threshold1.1_merge40_min60.bed",
                col.names = c('genome', 'peak_start', 'peak_end', 'name', 'score', 'strand', 'peakscore', 'peak_max')) %>%
                select(name, peak_start, peak_end, score) %>% mutate(peak_length = peak_end-peak_start, condition = 'LB')

M9_sharing <- read.table("../../processed_data/frag/Unshared_M9_peaks.bed", 
            col.names = c('genome', 'peak_start', 'peak_end', 'name', 'score', 'strand', 'peakscore', 'peak_max')) %>%
                select(name) %>% mutate(shared = 'Unique') %>% left_join(M9_peaks, ., by = 'name') %>% replace(., is.na(.),'Shared') 

LB_sharing <- read.table("../../processed_data/frag/Unshared_LB_peaks.bed", 
            col.names = c('genome', 'peak_start', 'peak_end', 'name', 'score', 'strand', 'peakscore', 'peak_max')) %>%
                select(name) %>% mutate(shared = 'Unique') %>% left_join(LB_peaks, ., by = 'name') %>% replace(., is.na(.),'Shared') 


d <- rbind(M9_sharing, LB_sharing) %>% ggplot(., aes(x=shared, y = score, fill = shared)) +
      stat_boxplot(geom='errorbar', linetype=1, width=0.25) +
      geom_boxplot(outlier.shape = NA) +
      scale_fill_manual(values = c('#e34f3c', '#2387bd')) +
      scale_y_log10() +
      #geom_jitter(alpha = .4) +
      geom_signif(comparisons = list(c("Unique", "Shared")), 
              y_position= 4.2, map_signif_level= function(p)sprintf("p = %.2g", p),
              test="wilcox.test", textsize = 5) +
      labs(x = 'Promoter Region', y='Region Score', title = '') +
      theme(legend.position = 'none') + facet_wrap(~condition) +
        theme(axis.text.x = element_text(size = 20), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20)) 


plot_grid(a,b,c,d, ncol = 2, labels = 'AUTO', label_size = 24)

ggsave('../../figs/supp_figs/S4_M9_Fragstats.png', units = 'in', width = 14.29, height = 10.51)

```




S5 - QC of scramble library

```{r}
scram_data <- read.table('../../processed_data/endo_scramble/endo_scramble_expression_formatted_std.txt', header = T)
neg <- subset(scram_data, grepl("neg_control", scram_data$name)) %>% subset(!grepl('scrambled', .$name))
threshold <- 1

cor <- cor(scram_data$RNA_exp_sum_1, scram_data$RNA_exp_sum_2)

a <- ggplot(scram_data, aes(RNA_exp_sum_1, RNA_exp_sum_2)) + 
  geom_point(alpha = .1) +
  scale_x_log10() + scale_y_log10() + annotation_logticks(sides= 'bl') +
  annotate("text", label=paste("italic(r)==",signif(cor, 3)),parse=TRUE, x=.5, y=15, size = 8) +
  xlab('Biological Replicate #1') + ylab('Biologial Replicate #2') +
  geom_point(data=neg,aes(RNA_exp_sum_1, RNA_exp_sum_2), color = "firebrick1") +
  annotate("text", label = "Negative\nControls", x = 10, y = .1, color = 'firebrick1', size = 8) +
  theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))

median_barcodes <- median(scram_data$num_barcodes_integrated)

b <- ggplot(scram_data, aes(x=num_barcodes_integrated)) +
      geom_histogram(color = 'black') +
      labs(x = "Number of barcodes",
            y = 'Counts') +
      geom_vline(xintercept = median_barcodes, color = 'red', linetype = 'dashed') +
      annotate('text', x = 28, y = 5000, 
                label = paste("Median =", median_barcodes, "barcodes"), size = 8) +
      theme(axis.text.x = element_text(size=24), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24))



plot_grid(a,b, ncol = 2, labels = 'AUTO', label_size = 24)

ggsave('../../figs/supp_figs/S5_Scramble QC_Fragstats.png', units = 'in', width = 12, height = 6)



```




```{r}

repressor_color <- '#000066'
activator_color <- '#FFD700'

data <-  read.table('../../processed_data/endo_scramble/endo_scramble_expression_formatted_std.txt', header = T)
ttest <- read.table('../../processed_data/endo_scramble/endo_scramble_ttests.txt', header = T)
data <- left_join(data, select(ttest, -tss_name), by = 'name')
scramble_merged <- read.table("../../processed_data/endo_scramble/scramble_merged_sites.txt", header = T)

scramble_merged_parts <- scramble_merged %>% 
    mutate(scramble_list = strsplit(scrambles, split = ',TSS')) %>% 
    unnest(scramble_list) %>% 
    mutate(scramble_part = gsub('TSS', '', scramble_list),
           scramble_part = paste0('TSS', scramble_part),
           scramble_part = gsub('TSSneg', 'neg', scramble_part)) %>% 
    select(-scramble_list) %>% 
    left_join(select(data, name, scramble_pos_rel_tss), by = c('scramble_part' = 'name'))
```

#Figure S6 - Global Analysis of Scrambles
```{r}

a <- data %>%  
    mutate("classification" = ifelse(relative_exp > 1, "Repressing", "Activating")) %>%
    filter(category != 'unscrambled') %>%
    ggplot(. , aes(relative_exp, fill = classification)) +
      geom_histogram(color = 'black', bins = 100) +
      scale_fill_manual(values = c('#FFD700', '#000066')) +
      labs(x = "Scramble relative expression",
           y = 'Counts') +
      scale_x_log10() +
      annotation_logticks(sides = 'b') +
      theme(axis.text.x = element_text(size = 20, angle = 45, hjust = 1), 
                axis.title.x = element_text(size = 20),
                axis.text.y = element_text(size = 20), 
                axis.title.y = element_text(size = 20), 
                legend.text=element_text(size=20),
                legend.position = 'top',
                legend.title = element_blank()) 

b <- scramble_merged %>% 
        mutate(length = scramble_end_pos - scramble_start_pos) %>% 
        mutate(site_type_mod = case_when(.$site_type == 'increased' ~ 'Activating',
                                     .$site_type == 'reduced' ~ 'Repressing')) %>%
        select(scramble_start_pos, scramble_end_pos,length, site_type_mod) %>%
        distinct() %>%
        ggplot(aes(length)) + 
        geom_histogram(aes(fill = site_type_mod), position = 'dodge', binwidth = 10, color = 'black') +
        scale_fill_manual(values = c(repressor_color, activator_color), na.value = 'gray65') +
        scale_x_continuous(breaks = seq(0, 100, 10)) +
        labs(fill = '', x = 'Merged scramble length (bp)', y = 'Counts') +
          theme(axis.text.x = element_text(size=20, angle = 45, hjust = 1), 
            axis.text.y = element_text(size=20),
            axis.title.x.bottom = element_text(size = 20),
            axis.title.y.left = element_text(size = 20),
            legend.text=element_text(size=20),
            legend.position = 'top')


c <- scramble_merged_parts %>% 
    mutate(length = scramble_end_pos - scramble_start_pos,
           length_bin = cut(length, breaks = seq(0, 100, 10)),
           site_type_mod = case_when(.$site_type == 'increased' ~ 'Activating',
                                 .$site_type == 'reduced' ~ 'Repressing')) %>%
    distinct() %>%
    left_join(select(data, name, relative_exp), by = c('scramble_part' = 'name')) %>%
    select(length, scramble_start_pos, scramble_end_pos, scrambles, length_bin, relative_exp, site_type_mod) %>% 
    distinct() %>% na.omit() %>%
    ggplot(aes(length_bin, relative_exp)) + 
    geom_boxplot(aes(fill = site_type_mod), position = 'dodge') +
    scale_y_log10() + annotation_logticks(sides = 'l') +
    scale_fill_manual(values = c(repressor_color, activator_color), na.value = 'gray65') + 
    labs(fill = '',
         x = 'Merged scramble length (bp)', 
         y = 'Relative expression') +
    theme(axis.text.x = element_text(size=20, angle = 45, hjust = 1), 
        axis.text.y = element_text(size=24),
        axis.title.x.bottom = element_text(size = 24),
        axis.title.y.left = element_text(size = 24),
      legend.text = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.position = 'top')



d <- scramble_merged_parts %>% 
    filter(site_type == 'reduced', !is.na(tf_concordance)) %>%
    ggplot(aes(scramble_pos_rel_tss)) + 
    geom_bar(aes(fill=tf_concordance)) +
    scale_x_continuous(breaks = c(-120,-80,-40,0)) +
    scale_fill_manual(values = c('#FFBC42', '#0496FF'), na.value = 'gray38') +
    labs(fill = 'TF concordance',
         x = 'Scramble position relative to TSS',
         title = 'Repressors') +
        theme(axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20),
        legend.position = 'top')

e <- scramble_merged_parts %>% 
    filter(site_type == 'increased', !is.na(tf_concordance)) %>%
    ggplot(aes(scramble_pos_rel_tss)) + 
    geom_bar(aes(fill=tf_concordance)) +
    scale_x_continuous(breaks = c(-120,-80,-40,0)) +
    scale_fill_manual(values = c('#FFBC42', '#0496FF'), na.value = 'gray38') +
    labs(fill = 'TF concordance',
         x = 'Scramble position relative to TSS',
         title = 'Activators') +
        theme(axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20),
        legend.position = 'top')



plot_grid(a,b,c,d,e, ncol = 3, labels = 'AUTO', label_size = 24)

ggsave('../../figs/supp_figs/S6_GlobalScrambles.png', units = 'in', width = 21.29, height = 10.51)

```


S7 - Random Promoters
```{r}


#How many active randoms?
act_rand <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'random' & active == 'active') %>% nrow()
#How many total randoms?
rand <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'random') %>% nrow()

#How many active pos_controls?
act_pos <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'pos_control' & active == 'active') %>% nrow()
#How many total positive controls?
pos <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'pos_control') %>% nrow()

#How many active randoms?
act_neg <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'neg_control' & active == 'active') %>% nrow()
#How many total randoms?
neg <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'neg_control') %>% nrow()

#How many active promoter tiles?
act_tile <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'tile' & active == 'active') %>% nrow()
#How many total promoter tiles?
tile <- read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt',
                        header = T) %>% filter(category == 'tile') %>% nrow()

read.table('../../processed_data/peak_tile/peak_tile_expression_formatted_std.txt', header = T) %>% 
    ggplot(aes(category, expn_med_fitted_scaled)) + 
    geom_boxplot(fill = 'dodgerblue', alpha = .4) +
    scale_y_log10(limits = c(.5, 500)) +
    annotation_logticks(sides = 'l') +
    geom_hline(yintercept = 1, linetype = 'dashed') +
    scale_x_discrete(limits = c('pos_control', 'tile', 'random', 'neg_control'),
                    labels = c('Positive\ncontrols', 'Promoter\nregion\ntiles', 'Random\n150mers', 'Negative\ncontrols')) +
    scale_fill_discrete(labels = c('Positive controls', 'Promoter region tiles', 'Random 150mers', 'Negative controls')) +
    labs(x = '', color = '', y = 'Expression') +
    annotate('text', x='random', y = 200, label = paste(paste(act_rand, '/', rand, sep = ''),
                                                        ' active', '\n(', signif(act_rand/rand,3)*100, '%)', sep = ''), size = 5) +
    annotate('text', x='pos_control', y = 200, label = paste(paste(act_pos, '/', pos, sep = ''),
                                                             ' active', '\n(', signif(act_pos/pos,3)*100, '%)', sep = ''), size = 5) +
    annotate('text', x='neg_control', y = 200, label = paste(paste(act_neg, '/', neg, sep = ''),
                                                             ' active', '\n(', signif(act_neg/neg,3)*100, '%)', sep = ''), size = 5) +
    annotate('text', x='tile', y = 200, label = paste(paste(act_tile, '/', tile, sep = ''),
                                                          ' active', '\n(', signif(act_tile/tile,3)*100, '%)', sep = ''), size = 5) +
    theme(axis.text.x = element_text(size=20), 
        axis.text.y = element_text(size=20),
        axis.title.x.bottom = element_text(size = 20),
        axis.title.y.left = element_text(size = 20),
        legend.position = 'none') 

    

ggsave('../../figs/supp_figs/S7_RandomActivity.png', units = 'in', width = 9.25, height = 5.25)


```











